---
title: "Learning Shiny"
output: github_document
---

```{r setup, include=FALSE}
# save the built-in output hook
hook_output <- knitr::knit_hooks$get("output")

# set a new output hook to truncate text output
knitr::knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x <- xfun::split_lines(x)
    if (length(x) > n) {
      # truncate the output
      x <- c(head(x, n-1), 
             "....", 
             tail(x, 2))
    }
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to Learning Shiny materials. This repository contains all necessary materials for the class  Introduction to Shiny – interactive apps with R held on 27/04/2022 in Big Data Institute, Oxford. 
 
## Requirements

* RStudio: download it [here](https://www.rstudio.com/products/rstudio/download/)  
* R (required >= 3.6, recommedned >= 4): download it [here](https://www.r-project.org/) 
* various R packages (for required packages see [setup](#setup))

You can look for instructions for installing RStudio and R here: [Installing R and RStudio](https://techvidvan.com/tutorials/install-r/). However, note that this workshop assumes certain knowledge of both R and RStudio. You don't have to be a pro but if you haven't used R & RStudio before you might be lost during this workshop. In such case it might be wise to go through some materials beforehand, for example from [here](https://ngschool.eu/ngseminars2020/) or [here](https://www.coursera.org/learn/r-programming). 


### Setup

To prepare for class please clone this repository and run the preparation script.

If the script executes properly the following messages should be printed.

```{r, include=FALSE, message=FALSE, error=FALSE, warning=FALSE, echo=FALSE}
source("scripts/prep_help.R")
```

```{r, out.lines = 4}
source("scripts/prep_help.R")
```

*Note:* The repo might change since last time you clone it, so please update it the evening before and check the above setup then.

