---
title: "Presenting gapminder dataset"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task outline

Your task is to prepare a Shiny App, or a dashboard to present and allow exploration of the [`Gapminder` dataset](https://www.gapminder.org/data/documentation/). In here I am using an excerpt from the data, imported from [`gapminder` r package](https://github.com/jennybc/gapminder) created by Jenny Bryan.

# About the gapminder dataset

Gapminder data is known from, among others, this amazing TED Talk: [The best stats you've ever seen | Hans Rosling](https://www.youtube.com/watch?v=hVimVzgtD6w). In the talk Hans Rosling presented the power of data visualisation to explore the world. If you are interested in data visualisation and haven't heard his talks, do look them up. 

`gapminder` dataset from R package with the same name contains 1704 observations with 6 variables:

* country
* continent
* year (every 5 years from 1952 to 2007)
* pop (total population)
* gdpPercap (GDP per capita), and
* lifeExp (life expectancy)

There are 12 observations per each country, with 142 countries. More information on the dataset can be found in the [Exploring the data](#exploring-the-data) section and in the documentation of the [`gapminder` package](https://github.com/jennybc/gapminder).

If you are curious and want to explore more, be sure to check out full dataset and other datasets gathered by the [Gapminder team](https://www.gapminder.org/data/documentation/). However, the goal of this exercise is to practice learning `Shiny` hence why, at least at the beginning, sticking with the excerpt will be best.


# Exploring the data

```{r, message=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(gapminder)
library(gtsummary)
thematic::thematic_rmd()
theme_update(legend.position = "bottom", legend.box = "vertical")
```

```{r, data}
gapminder %>%
  rename(Country = country, Continent = continent, Year = year,
         `Life expectancy at birth` = lifeExp, 
         `Total population` = pop,
         `GDP per capita in Int Dollars` = gdpPercap) %>%
  slice_sample(n = 10) %>%
  DT::datatable()
```

```{r, eval=FALSE}
gapminder %>%
  select(-country) %>%
  tbl_summary(label = list(continent ~ "Continent", year ~ "Year",
                           lifeExp ~ "Life Expectancy", pop ~ "Population",
                           gdpPercap ~ "GDP per Capita"),
              type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c(
                "{median} ({p25}, {p75})",
                "{min}, {max}")) %>%
  bold_labels() %>%
  modify_header(label ~ "**Variable**") 
```

```{r gt_summary, echo=FALSE, out.width="50%"}
gapminder %>%
  select(-country) %>%
  tbl_summary(label = list(continent ~ "Continent", year ~ "Year",
                           lifeExp ~ "Life Expectancy", pop ~ "Population",
                           gdpPercap ~ "GDP per Capita"),
              type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c(
                "{median} ({p25}, {p75})",
                "{min}, {max}")) %>%
  bold_labels() %>%
  modify_header(label ~ "**Variable**") %>%
  as_gt() %>%
  gt::gtsave("README_files/gt_summary.png")
```

# Some tips, ideas and code

First of all, you can show the relationship between life expectancy and country's wealth across the years. 

```{r, bubble_plot}
gapminder %>%
  ggplot(aes(gdpPercap, lifeExp, color = continent, size = pop)) +
  geom_point() +
  facet_wrap(~year) +
  scale_x_log10() + 
  labs(title = "Relationship between life expectancy and country's wealth",
       x = "GDP per capita [Int'l$]",
       y = "Life expectancy", 
       color = "Continent", size = "Population") +
  theme(legend.position = "bottom", 
        legend.box = "vertical")
```

For example, you might want to create a reactive plot, where the user will be able to choose a country to highlight and a year to show on the plot. 

```{r, select_year_county_plot}
selected_countries <-
  sample(unique(gapminder$country), 7)
selected_year <- "2002"

gapminder %>%
  mutate(selected = country %in% selected_countries) %>%
  filter(year == selected_year) %>%
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point(color = "grey60", aes(size = pop)) +
  geom_point(data = . %>% filter(selected), aes(size = pop,
                                                color = continent)) +
  ggrepel::geom_label_repel(data = . %>% filter(selected), 
                            aes(label = country)) +
  scale_x_log10() + 
  labs(title = "Relationship between life expectancy and country's wealth",
       subtitle = glue::glue("Data for the year {selected_year}"),
       x = "GDP per capita [Int'l$]",
       y = "Life expectancy", 
       color = "Continent", size = "Population") 
```

Another idea for visualising the data is to look at the change of life expectancy. 

```{r, life_expectancy_plot}
gapminder %>%
  ggplot(aes(year, lifeExp, group = country, color = continent)) +
  geom_line() +
  facet_wrap(~continent) +
  scale_color_manual(values = continent_colors) + 
  labs(title = "Change of life expectancy across the years",
       x = "Year",
       y = "Life expectancy", 
       color = "Continent") +
  scale_x_continuous(breaks = unique(gapminder$year)[c(TRUE, FALSE, FALSE)])
```

Or to look up wealth per continent. 

```{r, gdp_boxplot}
gapminder %>%
  filter(year == selected_year) %>%
  group_by(continent) %>%
  mutate(continent_median = median(gdpPercap)) %>%
  ggplot(aes(x = reorder(continent, continent_median), 
             y = gdpPercap, color = continent)) +
  gghalves::geom_half_point() +
  gghalves::geom_half_boxplot(outlier.shape = NA) +
  # if gghalves not available
  # geom_jitter() +
  # geom_boxplot(width = .1, outlier.shape = NA, 
  #              alpha = 0.4, color = "grey60") + 
  scale_color_manual(values = continent_colors) + 
  labs(title = glue::glue("Wealth across continents in {selected_year}"),
       y = "GDP per capita [Int'l$]", 
       color = "Continent") +
  scale_y_log10() +
  theme(axis.title.x = element_blank(), 
        legend.position = "none")
```

# Your turn! 

Those are just some ideas, now it's time for you to create your own Shiny App and explore the dataset even more! Good luck. :)